Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-26 VMCAI-revised-version/final.tex
up

\noindent
{\bf Our main contributions}
\begin{itemize}
\item Surprinsingly, the general undecidability statements about resilience were not known neither proved. We show that resilience and {state-resilience} (i.e. $X$ a singleton) problems are both undecidable for WSTS with strong compatibility and for $\Safe$ upward-closed or downward-closed. 
%Moreover, {state-resilience},
%{bounded-state-resilience} and
%{$k$-state-resilience}
%are undecidable for reset-VASS (that are strongly compatible $\pred$-basis-effective WSTS) when
%$\Safe$ is upward-closed. 
%			We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
\item The three resilience problems are decidable for post-ideal-effective Well Behaved Transition Systems (a generalisation of WSTS where the quasi-ordering is not necessarly well founded) with strong (upward) compatibility and when $\Safe$ is upward-closed (Theorem \ref{down-up})
%
\item  Resilience is also decidable, when 
$\Safe$ is upward-closed, for effective pred-basis WSTS with strong (upward) compatibility (Theorem \ref{xxx}).
%
%    and
%the additional hypothesis that
%for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
%
%				and for post-ideal-effective downward-WSTS ????
%%%
%
\item We clarify the different effectiness conditions on WBTS and WSTS that allow the decidability of the six resilience.
%
\item We generalize  the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} ; we also show that relaxing some hypothesis leads to undecidability.
%
%			à utiliser dans le texte ???
%
%We first remark that the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} can be generalized by relaxing the  strong compatibility hypothesis. We show that removing the effective 
%$\mathop{\uparrow}$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, many WSTS (like reset-VASS) don't enjoy this property.

@@

Most of our results about the different models and different types of resilience are in Figure YYYY.

@@

\alain{je pense qu'il faut faire deux tableaux pour que ce soit lisible. attention, la décidabilité n'est pas encore indiquée}

@@

@@

In order to construct algorithms, we also require effective hypotheses about the computations of certain sets of predecessors and successors. More precisely, we say that an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ is 
\begin{enumerate}
%\item {\em effective} if there exists a pair of algorithms
%($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
%$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
%$M_\leq$ the ordering relation “$\leq$”.
%
\item {\em $\post$-effective} if $\mathscr{S}$ is effective, and if there
exists an algorithm that computes $|\post(s)| \in \N \cup \{\infty\}$ 
on input $s \in S$.
%
%\item {\em ideally-effective} \cite{BFM-ic17} if the function mapping the encoding of a state $s$
%to the encoding of the ideal $\downarrow s$ is computable.
%
\item {\em $\post$-ideal-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any ideal $I \in Ideals(S)$ and returning $\mathop{\downarrow} \post(I)$, expressed as a finite union of (maximal) ideals.
%\item {\em completion-$\post$-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is $\post$-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
%
\item {\em $\pred$-basis-effective} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \pred(\mathop{\uparrow} s)$.
%
%and there exist an algorithm $M_\downarrow$ that computes 
% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\mathop{\uparrow}^C}$ that computes the ideal decomposition of $S \setminus \mathop{\uparrow} \{ s_1, s_2, \ldots, s_m\}$, on input
% $s_1, s_2, \ldots, s_m \in S$.
%
\item {\em $\post^*$-basis-effective} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \post^*(s)$.
%
\item {\em $\pred^*$-basis-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any ideal $I \in Ideals(S)$ and returning a finite basis of $\mathop{\downarrow} \pred^*(I)$.
\end{enumerate}

@@

	\vspace{-0.8cm}

@@

$\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility include for instance
Lossy Channel Systems with insertions~\cite{DBLP:journals/iandc/AbdullaJ96,DBLP:journals/iandc/CeceFI96}.

>>>>

\noindent
{\bf Our main contributions.}
\begin{itemize}
\item Surprinsingly, the general undecidability statements about resilience were not known neither proved. We show that resilience and {state-resilience} (i.e. $X$ a singleton) problems are both undecidable for WSTS with strong compatibility and for $\Safe$ upward-closed or downward-closed. 
%Moreover, {state-resilience},
%{bounded-state-resilience} and
%{$k$-state-resilience}
%are undecidable for reset-VASS (that are strongly compatible $\pred$-basis-effective WSTS) when
%$\Safe$ is upward-closed. 
%			We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
\item The three resilience problems are decidable for post-ideal-effective Well Behaved Transition Systems (a generalisation of WSTS where the quasi-ordering is not necessarly well founded) with strong (upward) compatibility and when $\Safe$ is upward-closed (Theorem \ref{down-up})
%
\item  Resilience is also decidable, when 
$\Safe$ is upward-closed, for effective pred-basis WSTS with strong (upward) compatibility (Theorem \ref{xxx}).
%
%    and
%the additional hypothesis that
%for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
%
%				and for post-ideal-effective downward-WSTS ????
%%%
%
\item We clarify the different effectiness conditions on WBTS and WSTS that allow the decidability of the six resilience.
%
\item We generalize  the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} ; we also show that relaxing some hypothesis leads to undecidability.
%
%			à utiliser dans le texte ???
%
%We first remark that the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} can be generalized by relaxing the  strong compatibility hypothesis. We show that removing the effective 
%$\mathop{\uparrow}$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, many WSTS (like reset-VASS) don't enjoy this property.

@@

Most of our results about the different models and different types of resilience are in the following tables. The different notions of effectivity are detailed on page~\pageref{on effectivity}.

@@

@@

% \alain{je pense qu'il faut faire deux tableaux pour que ce soit lisible. attention, la décidabilité n'est pas encore indiquée}

% Section 2 introduit les notions, Section 3 et 4 les résultats de résilience et state résilience pour les WBTS, Section 5 sur les VASS et variantes
Section~2 introduces the necessary definitions of WSTS and variations as well as the different notions of effectivity. Section~3 and 4 are concerned with the general framework of WSTS and contain our results on the three resilience problems
and the three state-resilience problems respectively. Lastly, Section~5 concerns itself with the resilience problems for VASS and variations of VASS.

@@

In order to construct algorithms, we also require effective hypotheses about the computations of certain sets of predecessors and successors. More precisely, we say that an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ is 
\begin{enumerate}\label{on effectivity}
%\item {\em effective} if there exists a pair of algorithms
%($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
%$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
%$M_\leq$ the ordering relation “$\leq$”.
%
\item {\em $\post$-effective} if $\mathscr{S}$ is effective, and if there
exists an algorithm that computes $|\post(s)| \in \N \cup \{\infty\}$ 
on input $s \in S$.
%
%\item {\em ideally-effective} \cite{BFM-ic17} if the function mapping the encoding of a state $s$
%to the encoding of the ideal $\downarrow s$ is computable.
%
\item {\em $\post$-ideal-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any ideal $I \in Ideals(S)$ and returning $\mathop{\downarrow} \post(I)$, expressed as a finite union of (maximal) ideals.
%\item {\em completion-$\post$-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is $\post$-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
%
\item {\em $\pred$-basis-effective} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \pred(\mathop{\uparrow} s)$.
%
%and there exist an algorithm $M_\downarrow$ that computes 
% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\mathop{\uparrow}^C}$ that computes the ideal decomposition of $S \setminus \mathop{\uparrow} \{ s_1, s_2, \ldots, s_m\}$, on input
% $s_1, s_2, \ldots, s_m \in S$.
%
\item {\em $\post^*$-basis-effective} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \post^*(s)$.
%
\item {\em $\pred^*$-basis-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any ideal $I \in Ideals(S)$ and returning a finite basis of $\mathop{\downarrow} \pred^*(I)$.
\end{enumerate}

@@

	\vspace{-0.9cm}

@@

$\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility include for instance
Lossy Channel Systems with insertions~\cite{DBLP:journals/iandc/AbdullaJ96%,DBLP:journals/iandc/CeceFI96
%
%% GOTO bibliographic reference when doing the bibtex final.aux did not work
}.
